---
title: "TidyTemplate"
date: 2023-04-26
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(janitor) #clean names
library(stringr) #leading zeros
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-04-25")

winners <- janitor::clean_names(tt$winners) #col names all lowercase
london <- janitor::clean_names(tt$london_marathon)
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

glimpse(winners) # woohoo time is already structured as time

glimpse(london)

#lots of Na in raised category
summary(london$raised)
length(london$raised)

unique(london$official_charity)
```

Winners:
- compare times for different categories of runner over time
- what athletes have to most entries, who over longest span of time
- a bunch of circles w area equivalent to nationality entered, animate to change by year
- how times change each year per nationality, could scrape weather conditions to see if nationality performance(how to measure?) and temp have correlation

## Scatterplots
```{r}
#restructure wheelchair and non-wheelchair side by side
lev <- c("Men", "Women", "Wheelchair Men", "Wheelchair Women")
winners$category <- factor(winners$category, levels = lev)

# 
winners |>
  ggplot() +
  geom_point(aes(x = year, y = time, color = category))

winners |>
  ggplot() +
  geom_point(aes(x = year, y = time)) +
  facet_wrap(~category)

# compare men to women, wheelchair to non-wheelchair
winners |>
  filter(grepl("Wheelchair", category)) |>
  ggplot() +
  geom_point(aes(x = year, y = time, color = category)) +
  scale_fill_hue(guide = guide_legend(reverse = TRUE))

winners |>
  filter(!grepl("Wheelchair", category)) |>
  ggplot() +
  geom_point(aes(x = year, y = time, color = category))

```

```{r testing}
bp <- ggplot(data=PlantGrowth, aes(x=group, y=weight, color = group)) + geom_point()
bp

bp + guides(fill = guide_legend(reverse=TRUE))

bp + scale_fill_discrete(guide = guide_legend(reverse=TRUE))

bp + scale_fill_discrete(breaks = rev(levels(PlantGrowth$group)))
```

London:
- applicants over years,
- acceptance rates by year
- finish rate per year
- 
## Lollipop Chart

manipulated lollipop code originally from @nrennie

```{r}
# lolli <- 
  london |>
  select(year, applicants) |>
  drop_na() |>
  mutate(yr = year %% 100) |> #get last two numbers of each year
  mutate(yr = str_pad(yr, 2, pad = "0")) |> #leading zeros for 2000s

  ggplot()+
  
  ## sticks
  geom_segment( 
    aes(x = year, xend = year,
        y = 0, yend = applicants)) +
  
  ## candies
  geom_point( #outline
    aes(x = year, y = applicants),
    colour = "white" ,size =9) +
  geom_point( # center color
    aes(x = year, y = applicants),
    colour = "red" ,size = 8) +
  
  ## year labels in candy
  geom_text(
    aes(x = year, y = applicants + 200,
        label = paste0(yr)),
    colour = "black", size = 4) +
  
  ## participant counts
  geom_text(
    aes(x = year, y = applicants + 100000,
        label = paste0(format(applicants, big.mark=","))),
    colour = "black", size = 4, angle = 90) +
  
  ## labels
  labs(title = "London Marathon",tag = "london marathon data, very cool") + 
  scale_x_continuous(limits = c(1980.5, 2020.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 550000), expand = c(0, 0)) +
  theme_void()

  theme(
    plot.title = element_text(
      lineheight = 0.4,
      colour = "black",
      hjust = 0,
      vjust = -8,
      face = "bold",
      size = 60,
      margin = margin(b = 5, t = 60, l = 15)),
    plot.title.position = "plot",
    plot.tag.position = c(0.03, 0.5),
    plot.tag = element_textbox_simple(
      lineheight = 0.45,
      colour = "black",
      width = 0.9,
      family = "Commissioner",
      hjust = 0,
      halign = 0,
      size = 24,
      margin = margin(b = 10, t = 5)
    ),
    plot.margin = margin(-80, 0, 0, 0)
    )

background <- 
  data.frame(x = 1:100, y = rep(100, 1000)) |>
  ggplot() +
  geom_area(
    mapping = aes(x = x, y = y),
    fill = "gray",
    alpha = 0.3
  ) +
  coord_cartesian(expand = FALSE) +
  theme_void() +
  theme(plot.margin = margin(0, 0, 0, 0))

background + patchwork::inset_element(lolli, 0, 0, 1, 1,
                     align_to = "full",
                     clip = TRUE)
```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
