---
title: "Portal Project"
date: 2023-05-02
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-05-02")

# plots <- tt$plots
species <- tt$species
surveys <- tt$surveys

```


# Readme

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

```{r Glimpse}

tt %>% 
  map(glimpse)

length(unique(surveys$plot))
length(unique(tt$plots$plot))

surveys |>
  select(plot, treatment) |>
  distinct(plot, treatment) |>
  arrange(plot)
tt$plots
# everything in plots is already in surveys


glimpse(surveys)
unique(surveys$species)

surveys |>
  count(tag) |>
  arrange(-n)

#are there na entries in surveys$species
surveys |>
  filter (is.na(species)) |>
  ncol() # there are 22
```

# Wrangle

```{r Wrangle}
species_key <- species |>
  select(species, commonname) |>
  filter(!is.na(species))

#count how many times each species was seen each month
temp <- surveys |>
  left_join(species_key, 
            by = "species") |>
  select(commonname, month, year) |>
  filter(!is.na(commonname)) |>
  count(year, commonname)
  
  
ggplot(data = temp, aes(x = year, y = n)) +
  geom_smooth() +
  facet_wrap(~commonname)


```

## Copy @mvbloois

```{r}
library(tidyverse)
library(ggridges)
library(ggthemes)
library(ggtext)
library(glue)
library(janitor)
library(showtext)

font_add_google("Fira Sans", "font")
showtext_auto()

bg <- "#EFEFEF"
fg <- "#123456"

## plot ------
plots <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-02/plots.csv')
species <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-02/species.csv')
surveys <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-02/surveys.csv')

df <- surveys %>% 
  filter(! is.na(species)) %>% 
  inner_join(
    species,
    by = "species" #removed join_by(species)
  ) %>% 
  mutate(
    colour = fg,
    label = glue("<b style='color:{colour}'>{commonname}</b><br/><i style='color:{colour};font-size:20px;'>{scientificname}</i>")) %>% 
  select(species, label, censusdate)
  
df %>% 
  ggplot(aes(x = censusdate, y = fct_rev(label), fill = species)) +
  geom_density_ridges(scale = 2.5, size = 0.25, rel_min_height = .01,
                      alpha = 0.5) +
  #scale_x_date(date_breaks = c("1980","1990","2000","2010","2020")) +
  scale_fill_tableau(palette = "Tableau 20") + #fill colors w "Tableau 20" color pallet
  labs(
    x = NULL,
    y = NULL,
    title = "Of mice and rats",
    subtitle = str_wrap("The Portal Project is a long-term ecological study being conducted near Portal, AZ. Since 1977, the site has been used to study the interactions among rodents, ants and plants and their respective responses to climate. To study the interactions among organisms, they experimentally manipulate access to 24 study plots.", 80), #str_wrap takes text arg and len of text arg
    caption = "Data: Portal Project" #subtitle takes caption arg
  ) +
  guides(
    fill = "none" #no guide or legend on the side
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "font",
                        colour = fg,
                        size = 30),
    axis.text.y = element_markdown(lineheight = 0.2, #why element_markdown for y and element_text for x
                                   vjust = 0.1),
    axis.text.x = element_text(size = 40,
                               colour = fg),
    panel.grid.major.y = element_line(colour = "grey90"),
    plot.title.position = "plot",
    plot.title = element_text(size = 80),
    plot.subtitle = element_text(face = "italic",
                                 size = 30,
                                 lineheight = 0.3,
                                 margin = margin(5,10,15,10)),
    plot.margin = margin(20,20,20,20)
  )


ggsave("./2023/2023-18/portal-project.png",
       height = 8,
       width = 6,
       bg = bg)
system("open ./2023/2023-18/portal-project.png")

```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}


  
```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
